---
output: pdf_document
---
Coursera Satistical Inference - Course Project PART 2
======================================================
Author: Dino Chioda
Date: May 23, 2015

# Overview
In this project, which is the second part of the Course Project for the Statistical Inference course, we're going to analyze the ToothGrowth data in the R datasets package. We will provide a basic summary of the data ans use confidence intervals to test hypotheses for tooth growth by supplement and dose.

## Question 1: Load the ToothGrowth data and perform some basic exploratory data analyses
```{r echo=TRUE}
# Load the data
library(datasets)
data(ToothGrowth)

# Look at the data
nrow(ToothGrowth)
str(ToothGrowth)
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
str(ToothGrowth)
```


## Question 2: Provide a baisc summary of the data

Using the `summary()` R function on the ToothGrowth dataset yields: 
```{r echo=FALSE}
summary(ToothGrowth)
```

Let's generate some basic box plots to gain a better understand of how the data behaves:
```{r}
library(ggplot2)
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose)) +
        ggtitle("Tooth Length by Dosage")
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp)) + 
        ggtitle("Tooth Length by Supplement")
```

## Question 3: Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.

Consider the following hypothesis:

$H_0$: Supplements have no effect on tooth length

$H_A$: Supplements have an effect on tooth length

We will use a t-test to assess that the $H_0$ is true.
```{r}
t.test(len~supp, ToothGrowth)
```
The p-value of the t-test is **0.06**. Therefore, we cannot reject the null hypothesis.

Consider the following hypothesis:

$H_0$: Increasing the dosage has no effect on tooth length

$H_A$: Increasing the dosage has an effect on tooth length

Since we have 3 distinct dosages, we have 3 possible dosage pairs to consider. Therefore, we will perform a t-test on each of the dosage pairings. Let's begin by subsetting by dosage pairings: 0.5 & 1.0, 0.5 & 2.0, 1.0 & 2.0:
```{r}
ToothGrowth.doses_0.5_1.0 <- subset (ToothGrowth, dose %in% c(0.5, 1.0)) 
ToothGrowth.doses_0.5_2.0 <- subset (ToothGrowth, dose %in% c(0.5, 2.0)) 
ToothGrowth.doses_1.0_2.0 <- subset (ToothGrowth, dose %in% c(1.0, 2.0))
```

Let's perform the 3 t-tests to validate $H_0$:
```{r}
t.test(len ~ dose, data = ToothGrowth.doses_0.5_1.0)
t.test(len ~ dose, data = ToothGrowth.doses_0.5_2.0)
t.test(len ~ dose, data = ToothGrowth.doses_1.0_2.0)
```

Since all 3 p-values are less than 0.05 and the confidence intervals do not contain 0, we can reject $H_0$ and state that increasing the dosage leads to increased tooth length.

## Question 4: State you conclusions and the assumptions needed for your conclusions.

###Conclusions
1. Supplement type does not seem to have an effect on tooth length.
2. Increasing dosage levels does appear to lead to increased tooth length.

### Assumptions
1. The sample population of guinea pigs is representative of the entire population of guinea pigs tested.
2. The sample population of guinea pigs were randomly selected.
3. All dosage pair subset considered have the same variance. 
